---
title: "cm008 Exercises: Fix the Plots"
author: Hannah McSorley
output: 
  html_document:
    keep_md: true
    theme: paper
---

In this worksheet, we'll be looking at some erroneous plots and fixing them. 

I think you might not have these two packages installed:

```
install.packages("ggridges")
install.packages("scales")
```

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(gapminder)
library(ggridges)
library(scales)
```

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

```{r try some plots}

# plot points only
ggplot(gapminder, aes(continent, lifeExp)) +
    geom_point()

# try with jitter (horizontal)
ggplot(gapminder, aes(continent, lifeExp)) +
    geom_jitter()

# NEAT! It spreads overlapping data for visibility but does not alter the values

# try with jitter and alpha transparency 
ggplot(gapminder, aes(continent, lifeExp)) +
    geom_jitter(alpha = 0.5)  # alpha has to go into the visual call (not aes)


# mountains of density?
ggplot(gapminder, aes(lifeExp, continent)) +
  ggridges::geom_density_ridges() +
  geom_rug()  # shows where the measurements were taken


# line vs path
tsibble::as_tsibble(co2) %>% 
  rename(yearmonth = index,
         conc = value) %>% 
  mutate(month = lubridate::month(yearmonth, label = TRUE),
         year  = lubridate::year(yearmonth)) %>% 
  ggplot(aes(month, conc)) +
  geom_line(aes(group = year), alpha = 0.5) +
  ylab("CO2 Concentration")

tsibble::as_tsibble(co2) %>% 
  rename(yearmonth = index,
         conc = value) %>% 
  mutate(month = lubridate::month(yearmonth, label = TRUE),
         year  = lubridate::year(yearmonth)) %>% 
  ggplot(aes(month, conc)) +
  geom_path(aes(group = year), alpha = 0.5) +
  ylab("CO2 Concentration")
```

## Exercise 1: Overlapping Points

After fixing the error, fix the overlapping problem in the following plot (attribution: ["R for data science"](https://r4ds.had.co.nz/data-visualisation.html)).

```{r, fig.width = 5, fig.height = 2}
head(mpg)  # what is it?

ggplot(mpg, aes(cty, hwy)) +  # don't pipe, use plus in ggplot2
  geom_jitter(alpha = 0.5) +  # jitter to spread points, alpha for transparecy
  geom_smooth(method = "lm") +
  theme_bw()
  
# note that geom_smooth() applies to the data, not the jitter
# jitter is just a visual trick for clarifying
```


## Exercise 2: Line for each Country

Fix this plot so that it shows life expectancy over time _for each country_. Notice that `ggplot2` ignores the grouping of a tibble!

```{r, fig.width = 5, fig.height = 2}
gapminder %>% 
  #group_by(country) %>%  # ggplot is ignoring this grouping
  ggplot(aes(year, lifeExp, group = country, colour = continent)) +
  geom_line(alpha = 0.35) +
  theme_bw()

# try it with Rwanda highlighted
gapminder %>% 
  #group_by(country) %>%  # ggplot is ignoring this grouping
  ggplot(aes(year, lifeExp, group = country, colour = country == "Rwanda")) +
  geom_line() +
  theme_bw()  +
  scale_colour_discrete("Country", labels = c("Other", "Rwanda"))
```


## Exercise 3: More gdpPercap vs lifeExp

### 3(a) Facets

- Change the x-axis text to be in "comma format" with `scales::comma_format()`.
- Separate each continent into sub-panels.

```{r, fig.width = 5, fig.height = 2}
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point(alpha = 0.2) +
  facet_wrap(~ continent) +
  scale_x_log10(labels = scales::comma_format())  # labels changes from sci.notation
```

### 3(b) Bubble Plot

- Put the plots in one row, and free up the axes.
- Make a bubble plot by making the size of the points proportional to population. 
  - Try adding a `scale_size_area()` layer too (could also try `scale_radius()`).
- Use `shape=21` to distinguish between `fill` (interior) and `colour` (exterior). 

```{r, fig.width = 5, fig.height = 2}
gapminder %>% 
  filter(continent != "Oceania") %>% 
  ggplot(aes(gdpPercap, lifeExp, size = pop, fill = continent)) +  # 
  facet_wrap(~ continent, nrow = 1) +  # adding scales = free makes it less comparable
  geom_point(alpha = 0.5, shape = 21) +  # shape 21 is a hollow circle (use fill instead of colour)
  scale_x_log10(labels = scales::comma_format()) +
  scale_size_area()


# when you apply mapping to ggplot aes, it applies globally
# when you apply mapping to geom_"", it applies only to that layer

```

___- This is the class end point -___

A list of shapes can be found [at the bottom of the `scale_shape` documentation](https://ggplot2.tidyverse.org/reference/scale_shape.html).

### 3(c) Size "not working"

Instead of alpha transparency, suppose you're wanting to fix the overplotting issue by plotting small points. Why is this not working? Fix it.

```{r, fig.width = 5, fig.height = 2}
ggplot(gapminder) +
  geom_point(aes(gdpPercap, lifeExp, size = 0.1)) +
  scale_x_log10(labels = scales::dollar_format())
```



## Exercise 4: Walking caribou

The following mock data set marks the (x,y) position of a caribou at four time points. 

- Fix the plot below so that it shows the path of the caribou. 
- Add an arrow with `arrow = arrow()`.
- Add the `time` label with `geom_text()`.

```{r, fig.width = 5, fig.height = 2}
tribble(
  ~time, ~x, ~y,
  1, 0.3, 0.3,
  2, 0.8, 0.7,
  3, 0.5, 0.9,
  4, 0.4, 0.5
) %>% 
  ggplot(aes(x, y)) + 
  geom_line()
```

## Exercise 5: Life expectancies in Africa

### 5(a) Unhiding the data

Fix the plot so that you can actually see the data points. Be sure to solve the problem of overlapping text, without rotating the text.

```{r, fig.width = 5, fig.height = 2}
gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot(aes(country, lifeExp)) + 
  geom_point() +
  geom_boxplot()
```

### 5(b) Ridgeplots

We're starting with the same plot as above, but instead of the points + boxplot, try a ridge plot instead using `ggridges::geom_density_ridges()`, and adjust the `bandwidth`.

```{r, fig.width = 5, fig.height = 2}
gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot(aes(y = country, x = lifeExp)) + 
  ggridges::geom_density_ridges() + # you have to switch x and y from box to ridge
  

```

# ------------------------------

ENDED HERE OCT 03

# -----------------------------

## Exercise 6: Bar plot madness

### 6(a) Colour and stacking madness

- Change the following plot so that it shows _proportion_ on the y-axis, not count.
- Change the x-axis so that it doesn't appear to be continuous.
- Put the bars for transmission side-by-side with their own colour.
- Capitalize the legend title.

```{r, fig.width = 5, fig.height = 2}
mtcars %>% 
  mutate(transmission = if_else(am == 0, "automatic", "manual")) %>% 
  ggplot(aes(cyl)) +
  geom_bar(aes(colour = transmission))
```

### 6(b) Bar heights already calculated

Here's the number of people having a certain hair colour from a sample of 592 people:

```{r}
(hair <- as_tibble(HairEyeColor) %>% 
  count(Hair, wt = n))
```

Fix the following bar plot so that it shows these counts.

```{r}
ggplot(hair, aes(Hair, n)) +
  geom_bar()
```

## Exercise 7: Tiling

Here's the number of people having a certain hair and eye colour from a sample of 592 people:

```{r}
(hair_eye <- as_tibble(HairEyeColor) %>% 
  count(Hair, Eye, wt = n))
```

Fix the following plot so that it shows a filled-in square for each combination. 

```{r}
ggplot(hair_eye, aes(Hair, Eye)) +
  geom_point(aes(colour = n))
```

By the way, `geom_count()` is like `geom_bar()`: it counts the number of overlapping points.


## Additional take-home practice

If you'd like some practice, give these exercises a try

__Exercise 1__: Make a plot of `year` (x) vs `lifeExp` (y), with points coloured by continent. Then, to that same plot, fit a straight regression line to each continent, without the error bars. If you can, try piping the data frame into the `ggplot()` function.

__Exercise 2__: Repeat Exercise 1, but switch the _regression line_ and _geom\_point_ layers. How is this plot different from that of Exercise 1?

__Exercise 3__: Omit the `geom_point()` layer from either of the above two plots (it doesn't matter which). Does the line still show up, even though the data aren't shown? Why or why not?

__Exercise 4__: Make a plot of `year` (x) vs `lifeExp` (y), facetted by continent. Then, fit a smoother through the data for each continent, without the error bars. Choose a span that you feel is appropriate.

__Exercise 5__: Plot the population over time (year) using lines, so that each country has its own line. Colour by `gdpPercap`. Add alpha transparency to your liking. 

__Exercise 6__: Add points to the plot in Exercise 5.