---
title: "cm007 Exercises: Practice with `dplyr`"
author: "Hannah McSorley"
date: "24/09/2019"
output:
  html_document:
    keep_md: true
    theme: paper
---
<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tsibble))
```


This worksheet contains exercises aimed for practice with `dplyr`. 


1. (a) What's the minimum life expectancy for each continent and each year? (b) Add the corresponding country to the tibble, too. (c) Arrange by min life expectancy.

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(min_life = min(lifeExp), 
            country = country[lifeExp == min_life]) %>% ## call a new variable that is created in the same function call
  arrange(min_life)


# -----
# identify the groups with a nested function
groups(gapminder %>% group_by(continent, year))  
# identify the groups as a pipe
gapminder %>% 
  group_by(continent, year) %>% 
  groups()
```

2. Calculate the growth in population since the first year on record _for each country_ by rearranging the following lines, and filling in the `FILL_THIS_IN`. Here's another convenience function for you: `dplyr::first()`. 

```
mutate(rel_growth = FILL_THIS_IN)) %>% 
arrange(FILL_THIS_IN) %>% 
gapminder %>% 
DT::datatable()
group_by(country) %>% 
```

```{r}

gapminder %>% 
  group_by(country) %>% 
  arrange(year) %>% 
  mutate(rel_growth = pop - dplyr::first(pop)) %>% 
  DT::datatable()

```


3. Determine the country that experienced the sharpest 5-year drop in life expectancy, in each continent, sorted by the drop, by rearranging the following lines of code. Ensure there are no `NA`'s. Instead of using `lag()`, use the convenience function provided by the `tsibble` package, `tsibble::difference()`:

```
# drop_na() %>% 
# ungroup() %>% 
# arrange(year) %>% 
# filter(inc_life_exp == min(inc_life_exp)) %>% 
# gapminder %>% 
# mutate(inc_life_exp = FILL_THIS_IN) %>% 
# arrange(inc_life_exp) %>% 
# group_by(country) %>% 
# group_by(continent) %>% 
# knitr::kable()
```

```{r}
gapminder %>% 
  group_by(country) %>% 
  arrange(year) %>% 
  mutate(inc_life_exp = tsibble::difference(lifeExp)) %>% 
  drop_na() %>% 
  ungroup() %>% 
  group_by(continent) %>%  # you have to ungroup before you group by a new variable
  filter(inc_life_exp == min(inc_life_exp)) %>% 
  arrange(inc_life_exp) %>%  # arrange() works independent of groups
  knitr::kable()
```
