---
title: "STAT 545 markdown & tibble exploration"
author: "Hannah McSorley"
date: "10/09/2019"
output: 
    html_document:
      toc: true
      number_sections: true
      theme: cerulean
---

```{r load_packages, warning = FALSE, echo = FALSE}
# labelling your chunks will help you in future endeavours
# hide the warnings about which version each package was made under
# echo = FALSE means the code will NOT be shown in your knited output
library(gapminder)
library(tibble)
library(DT)
```
New! make a change (in GitHub) -- to explore conflicts between local and remote 

New! make a change now -- to explore conflicts between local and remote 

# Practice with some chunk work
Print out 'gapminder' data

```{r dataset}
gapminder  # life expectancy, GDP info for different parts of the world
# you'll be using this dataset throughout the course
```

Convert the 'mpcars' dataset to a tibble format and datatable format

```{r}
as_tibble(mtcars)  # tibble table has info about object class
mtcars  # basic dataframe format
datatable(mtcars)  # navigatable table, very tidy
```

make a comment inline
<!--this is an inline comment-->
note that it's helpful to use the syntax package::function (e.g. DT::datatable) because it helps to identify which package you're using without adding a comment (also helpful if you have several packages that might have some similar functions)
